Supplementary Material for `The response of trophic interaction networks to multiple stressors in a marine latitudinal gradient of the Southern Hemisphere'

### 2024-06-08 manuscript - 8Jun.Rmd
Add M&M details

Here we review for the first time stressor responses of marine food webs for areas located along a Southwest Atlantic to Antarctic gradient (45 - 78ºS, Figure 1). We considered proven and expected changes in food webs driven by global change-related stressors in selected areas along this large-scale latitudinal range.
The aim of this review is threefold: 1) describe the complexity and structure of marine food webs along the Southwest Atlantic to Antarctic axis from a network perspective; 2) identify the ongoing global change-related stressors for each marine ecosystem containing the food webs; and 3) review proven food web changes and elaborate on a series of alternative hypotheses on how the identified stressors might affect food web features (energy flow, stability and topology), combining information on node- and network-level properties. 
To identify the main stressors, we performed an exploratory and configurative review instead of a systematic review [@Haddaway2015], because our objective was to formulate evidenced-based hypotheses rather than providing an exhaustive list of species-level stressors for each ecosystem.
Finally, we suggest which additional data and analyses are needed to gain insights into the stressors’ effects on food web properties in the Southwest Atlantic to Antarctic region.

@@

Throughout this region, many studies address the ecology of predator-prey relationships but most of these studies are restricted to a few interacting species or certain components of the food web such as fish, crustaceans, and benthos (e.g., @Vinuesa2007, @Padovani2012, @Pasotti2015, @Riccialdelli2020, @Saunders2019, @Pasti2021, @Cordone2022). Only few studies consider the complexity of trophic interactions within highly resolved food webs containing hundreds of species, ranging from phytoplankton, zooplankton, benthos, fish, seabirds to whales (but see @Jacob2011, @Marina2018, @Funes2022, @Lopez-Lopez2022, @Rodriguez2022, @Marinainrev). Neglecting this complexity might lead to a misunderstanding of the structure and functioning of the ecosystems and ultimately reduce the ability to predict ecosystem responses to perturbations [@Montoya2009].

@@

In this review, we included six large marine areas in the Southern Hemisphere for which highly-resolved empirical food webs in terms of species diversity and trophic interactions have been previously described: (1) San Jorge Gulf (~45 - 47ºS, 65 - 68ºW) [@Funes2022], (2) Beagle Channel (~54ºS, 68ºW) [@Rodriguez2022], (3) Burdwood Bank (~54ºS, 59ºW) [@Marinainrev], (4) Scotia Sea (~48 - 58ºS, 50ºW) [@Lopez-Lopez2022], (5) Potter Cove (62ºS, 58ºW, Antarctica) [@Marina2018], and (6) Weddell Sea (74 - 78ºS, 30ºW, Antarctica) [@Jacob2011; @Marina2024] (Figure 1). The geographic locations of these marine biomes span from temperate Atlantic waters to Antarctic, and are exposed to global change stressors (e.g., sea warming, glacial retreat, fishery, pollution) [@Sage2020], of which some are locally more important, e.g., fisheries in the San Jorge Gulf or glacial retreat in the Antarctic.

@@

The Burdwood Bank food web comprises 379 nodes and 1788 interactions, with a connectance of 0.01, and an asymmetric degree distribution (i.e. most of the species have a relatively low number of interactions and few species concentrate most of them). Almost half of the consumers are omnivores (48%), and the network displays a small-world pattern [@Marinainrev] (Table 1). This network pattern implies a rapid spread of a perturbation (e.g., contaminant, population fluctuation, local extinction) throughout the food web due to a short distance among species but, at the same time, a greater resistance caused by the confinement of perturbations mainly within subnetworks as a result of a high clustering coefficient [@Bornatowski2017; @Dormann2017].

@@

Burdwood Bank         379         1788        0.01              2.99              2.52          0.49           @Marinainrev

>>>>

Here we review for the first time stressor responses of marine food webs for areas located along a Southwest Atlantic to Antarctic gradient (45 - 78ºS, Figure 1). We considered proven and expected changes in food webs driven by global change-related stressors in selected areas along this large-scale latitudinal range.
The aim of this review is threefold: 1) describe the complexity and structure of marine food webs along the Southwest Atlantic to Antarctic axis from a network perspective; 2) identify the ongoing global change-related stressors for each marine ecosystem containing the food webs; and 3) review proven food web changes and elaborate on a series of alternative hypotheses on how the identified stressors might affect food web features (energy flow, stability and topology), combining information on node- and network-level properties. 
To identify the main stressors, we performed an exploratory and configurative review instead of a systematic review [@Haddaway2015], where we considered more than 200 peer-reviewed articles retrieved mainly from Google Scholar using keywords combining location (i.e., "San Jorge Gulf", "Beagle Channel", "Burdwood Bank", "Scotia Sea", "Weddell Sea", "Potter Cove") and stressors (i.e., "climate change", "global warming", "fishery", "microplastics", etc). We performed this type of review because our objective was to formulate evidenced-based hypotheses rather than providing an exhaustive list of species-level stressors for each ecosystem.
Finally, we suggest which additional data and analyses are needed to gain insights into the stressors’ effects on food web properties in the Southwest Atlantic to Antarctic region.

@@

Throughout this region, many studies address the ecology of predator-prey relationships but most of these studies are restricted to a few interacting species or certain components of the food web such as fish, crustaceans, and benthos (e.g., @Vinuesa2007, @Padovani2012, @Pasotti2015, @Riccialdelli2020, @Saunders2019, @Pasti2021, @Cordone2022). Only few studies consider the complexity of trophic interactions within highly resolved food webs containing hundreds of species, ranging from phytoplankton, zooplankton, benthos, fish, seabirds to whales (but see @Jacob2011, @Marina2018, @Funes2022, @Lopez-Lopez2022, @Rodriguez2022, @Marina2024a). Neglecting this complexity might lead to a misunderstanding of the structure and functioning of the ecosystems and ultimately reduce the ability to predict ecosystem responses to perturbations [@Montoya2009].

@@

In this review, we included six large marine areas in the Southern Hemisphere for which highly-resolved empirical food webs in terms of species diversity and trophic interactions have been previously described: (1) San Jorge Gulf (~45 - 47ºS, 65 - 68ºW) [@Funes2022], (2) Beagle Channel (~54ºS, 68ºW) [@Rodriguez2022], (3) Burdwood Bank (~54ºS, 59ºW) [@Marina2024a], (4) Scotia Sea (~48 - 58ºS, 50ºW) [@Lopez-Lopez2022], (5) Potter Cove (62ºS, 58ºW, Antarctica) [@Marina2018], and (6) Weddell Sea (74 - 78ºS, 30ºW, Antarctica) [@Jacob2011; @Marina2024] (Figure 1). The geographic locations of these marine biomes span from temperate Atlantic waters to Antarctic, and are exposed to global change stressors (e.g., sea warming, glacial retreat, fishery, pollution) [@Sage2020], of which some are locally more important, e.g., fisheries in the San Jorge Gulf or glacial retreat in the Antarctic.

@@

The Burdwood Bank food web comprises 379 nodes and 1788 interactions, with a connectance of 0.01, and an asymmetric degree distribution (i.e. most of the species have a relatively low number of interactions and few species concentrate most of them). Almost half of the consumers are omnivores (48%), and the network displays a small-world pattern [@Marina2024a] (Table 1). This network pattern implies a rapid spread of a perturbation (e.g., contaminant, population fluctuation, local extinction) throughout the food web due to a short distance among species but, at the same time, a greater resistance caused by the confinement of perturbations mainly within subnetworks as a result of a high clustering coefficient [@Bornatowski2017; @Dormann2017].

@@

Burdwood Bank         379         1788        0.01              2.99              2.52          0.49           @Marina2024a
